\subsection{What is diKTat?}
DiKTat - is a coding convention (\url{https://github.com/cqfn/diKTat}) and a and a linter (code checker&fixer) with a set of rules that implement this code style for Kotlin language. Basically, it is a collection of Kotlin code style rules implemented as AST visitors on top of KTlint framework (\url{https://github.com/pinterest/ktlint}). Diktat warns and fixes code style errors and code smells based on configuration file. DiKTat is a highly configurable framework, that can be extended further by adding custom rules. It can be run as command line application or with maven or gradle plugins. In this paper, we will explain how DiKTat works, describes advantages and disadvantages and how it differs from other static analyzers.

\subsection{Why diKTat?}
So why did we decide to create diKTat? We looked at similar projects and realized that they have defects and their functionality does not give you a chance to implement modern configurable code style. We have understood that we simply will need to implement a lot of configurable rules that are missing in other static analyzers. But we decided not to create "yet another one framework with dirty hacks" and made up one's mind to implement our set of rules on the top of a Ktlint framework that we are using as a dependency in Diktat. That’s why we came to a conclusion that we need to create convenient and easy-to-use tool for developers. Why is it easy-to-use? First of all, diKTat has its own highly configurable ruleset. You just need to fill your own options on rules in ruleset or either use default one. Basically, ruleset is defined as a yml file with a description of each rule. Secondly, there are a lot of developers that use different tools for building projects. Most popular are Maven and Gradle. DiKTat supports these ones and it also has cli. Finally, each developer has their own codestyle and sometimes they don’t want static analyzers to trigger on some lines of code. In diKTat you can easily disable each inspection. Also diktat has unique inspections that are not present in other analyzers. It works in two modes:
\begin{itemize}
\item fixer - it fixes the code in the original file and prints warnings for inspections that cannot be autofixed (when linter does not understand developer's intention)
\item checker - it only outputs warnings for all issues that diktat can find
\end{itemize}